Generalists are the most urban-tolerant of birds: an analysis of ecological and life history traits using a novel continuous urbanization index

### 2018-09-03 ms.Rmd
revisions for next version of paper

author: |
  | Corey T. Callaghan^[Corresponding author: c.callaghan@unsw.edu.au] $^1$, Richard E. Major $^1$$^,$$^2$, John H. Wilshire $^1$, John M. Martin $^1$$^,$$^3$, Richard T. Kingsford $^1$, William K. Cornwell $^1$$^,$$^4$
  | $^1$Centre for Ecosystem Science, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia; $^2$Australian Museum Research Institute, Sydney, NSW, Australia; $^3$Royal Botanic Gardens and Domain Trust, Sydney, NSW, Australia; $^4$Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia
date: ""
output:
  rmarkdown::pdf_document:
    fig_caption: yes    
    includes:  
      in_header: preamble-latex.tex
fontsize: 12pt
link-citations: no
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/proceedings-of-the-royal-society-b.csl
bibliography: references.bib
---

@@

\newpage
# INTRODUCTION
Species' specialization, based on a suite of biological traits, can dictate population-level responses to changing environments. Generally, species with limited niche breadth experience higher extinction risk than those with wider niche breadth [@boulangeat2012niche]. Further, some organisms have traits that allow for adaptation of their niches with changing environments, leading to potential ecological inheritance [@odling2013niche]. Urbanization drastically changes natural habitats through destruction, fragmentation, and isolation [@marzluff2001restoration; @melles2003urban; @loram2007urban], often providing ecologically novel landscapes for species. Some species successfully adapt, colonize, and thrive in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation], while others, common in surrounding natural habitats, rarely occur in urban environments [@tratalos2007bird; @evans2011makes; @gatesire2014bird]. Urbanization selects for specific sets of species, globally threatening biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects], a major concern for conservation [@czech2000economic; @dearborn2010motivations]. It is important to identify which ecological and life history traits most associate with successful or unsuccessful species in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful; @mcdonnell2015adaptation]. This can improve understanding of relationships between biodiversity and urbanization, and guide protection of the species most at-risk of extinction, through habitat protection, management, and restoration.

Many ecological and life history traits are associated with bird species in urban environments, including degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution]; nesting substrate [@kark2007living; @croci2008does; @conole2011functional]; diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes]; range size [@jokimaki2000artificial; @chace2006urban; @croci2008does]; migratory status [@friesen1995effects]; niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes]; and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific] (see Appendix 1). In Jerusalem, urban bird species were best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while, throughout France and Switzerland, urban bird species were positively associated with forested habitats, sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does]. There are also contrasting results: residual brain size in birds is sometimes positively associated with urbanization [@maklakov2011brains; @moller2015brain] and sometimes not [@kark2007living; @evans2011makes]. Similarly, contrasting results have been reported for annual fecundity [cf., @croci2008does; @moller2009successful; @evans2011makes] and niche breadth [cf., @kark2007living; @evans2011makes].

Most analyses assign species to discrete a priori groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or strictly presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful]. These approaches severely limit the capacity to generalize, as it assumes grouped species respond equally to urbanization [@evans2011makes]. Terminology also varies among studies [cf., @kark2007living; @croci2008does; @geschkecompact], affecting interpretation [@mcdonnell2015adaptation; @fischer2015categorizing]. There is clearly a need for continuous measures of urbanization [@evans2011makes], where a species is assessed over its entire distribution in response to urban environments.

Most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does] or temporally limited to a single breeding season [@blair1996land] or even two-month survey period [@kark2007living]. Broad-scale empirical data allow hypotheses to be tested at large spatiotemporal scales. Citizen science can provide such data [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen], and can be used to describe patterns of abundance, distribution, and functional composition of birds in urban ecosystems [@mccaffrey2005using; @la2014role; @morelli2016evidence; @lepczyk2017global @callaghan2017assessing].

We aimed to determine what suite of life history traits predicted avian tolerance to urbanization at a continental scale. We predicted that different traits would positively and negatively associate with urbanization. For generality, our analyses had to (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3) include a broad-range of traits, and (4) employ a continuous measure of urbanization. We integrated a series of large datasets to establish a novel continuous urbanization index to test associations with ecological and life history traits influencing species' responses to urbanization. We first investigated whether our urbanization index was statistically independent in relation to phylogeny (i.e., phylogenetic signal - [@revell2008phylogenetic; @evans2011makes]), given our comprehensive sample size. We then developed models, including and excluding phylogenetic effects, to test for associations between our urbanization index and a suite of continuous and categorical predictor variables representing species' traits.

# METHODS
## Bird observation data
We collated bird observations throughout mainland Australia from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists. Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird], often comparable to trained observers [@callaghan2015efficacy]. We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018. This corresponded to the richest period of eBird data, minimizing undue leverage of mismatch between changes in response variables over which bird observations were collected.

We filtered bird observations [@callaghan2017assessing; @la2014role], only including observations on complete checklists and which were either travelling, random, stationary, area, or followed BirdLife Australia protocols [@barrett2003new] (see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types). We then filtered the checklists to those which travelled < 5 km or covered < 500 Ha, to ensure the highest quality data were used. Any checklists shared among multiple observers were subsampled, including only one checklist to avoid duplication. This provided 637,482 eligible checklists throughout Australia. Seabirds (e.g., gannets, petrels, shearwaters) were omitted from analyses as we did not expect any correlation with seabirds using urban areas. Further, only species with > 100 observations were analyzed. In total 580 species, most (~80%) of Australia's bird species, were candidates for analyses. We used differing taxonomic authorities (see Appendix 2).

## Urbanization index
We used VIIRS night-time lights [@elvidge2017viirs] as a continuous measure for urban habitat, measuring the electric lighting associated with human settlement (Fig. S1). This has filtered out background noise (e.g., fires, degraded data, and other light source contamination) to ensure validity of the data and its association with human settlement [@elvidge2017viirs]. For our checklists and their associated species, we used the spatial coordinates and Google Earth Engine [@gorelick2017google], to assign the average radiance value, calculated from the VIIRS layer. For each checklist, the average radiance was calculated within a 5 km buffer, to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols. Any negative radiance values were forced to 0.00001 [*sensu* @ou2015evaluation]. Every bird species subsequently had a distribution of average radiance values (Fig. 1), with the median of each species' distribution becoming the response variable. There was no definitive range for this index, dependent on a species' distribution of observations.

@@

## Statistical analyses
Before modelling of predictor variables (Table 1), we investigated multi-collinearity among continuous variables (Fig. S2), using the corrplot package [@corrplot2017]. We used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation among each other; all traits < 2 suggested negligible correlation (Supplementary Table 1). We used two independent linear modelling approaches: one without and one with phylogenetic effects. Phylogenetic effects are defined as the tendency of closely related organisms to have similar life history components due to their shared evolutionary history [@derrickson1988taxon; @mckitrick1993phylogenetic]. Both approaches were used because of shortcomings [@cadotte2017phylogenies] and importance [@losos2008phylogenetic; @graham2018phylogenetic] of incorporating evolutionary distinctness in ecological research.

Linear regression models were weighted [@solon2015weighting], using a source of error associated with our urbanization indices, which were log-transformed to meet model assumptions. This error was the proportion of the number of total observations for a species, divided by the number of its unique localities (Fig. S2 shows the correlation between the two). This weighted the model-fitting to those species with more observations in more unique locations, likely to be the most reliable urbanization indices. This approach also addressed biases in sampling of eBird data, where some birds have a high number of observations from a small number of locations such as known roosting sites [e.g., @callaghan2018avian]. All models and analyses were conducted in R [@r2017coreteam], relying on the tidyverse workflow [@wickham2017tidyverse]. Statistical significance was concluded as $\alpha$ < 0.05. Variability of means is reported as standard deviations.

@@

## Phylogenetic analyses
### Phylogenetic signal
There was a distinct relationship between the phylogenetic signal and our urbanization index (Fig. 3; Fig. S3; Table 2), significantly different from random for 4 out of 5 indices (Cmean, I, K.star, and Lambda). For example, the urbanization index was disproportionately important with respect to the phylogenetic tree (Fig. 3), evidenced by the highest urbanization indices clustered throughout the waterbird clade. Further, within a specific clade, there were some species with substantially higher urbanization indices (e.g., *Manorina melanocephala*, Noisy Miner, cf. honeyeaters). We generally found weaker signals for behavioral, compared with morphological traits. For behavioural traits, the estimates of phylogenetic signal were weaker than Brownian motion, but usually significant (Table 2). Additionally, there was a strong phylogenetic signal for body size and brain residual size (Table 2).

@@

# DISCUSSION
The most urban-tolerant species in our analyses, defined based on their urbanization indices, had large niche breadths, reflected in generalism. Conversely, more specialized species were the least urban-tolerant: specialized feeding strategies (i.e., granivorous and insectivorous diets) seldom occurred in urban environments. These results were robust across separate non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig. S4), indicating that urbanization is generally excluding specialist species [@sorace2009distribution; @concepcion2015impacts]. Further, the non-random distribution of the urbanization index across the phylogeny (Fig. 3 & Table 2) is consistent with reduced phylogenetic diversity in urban areas [@la2018phylogenetic]. Particular phylogenetic effects [@derrickson1988taxon; @mckitrick1993phylogenetic] probably allow specific subsets of species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].

@@

In conclusion, our novel continuous urbanization index showed relationships between ecological and life history traits with avian tolerance to urban environments. Improving understanding of the temporal nature of these relationships will be important, but we highlight the importance of considering these questions using a continuous measure of urbanization [@evans2011makes] as well as using a phylogenetically informed approach. It is the generalist species [@shochat2006patterns], with high clutch sizes and large residual brain sizes, which are best prepared to forage and survive in the ecologically novel aspects of urban environments.

# Data, code and materials
This work relied heavily on open-source, fully accessible datasets.
The necessary portions of those datasets, along with code to reproduce the analyses are available at
this github repository (https://github.com/cornwell-lab-unsw/bird_urbanness), and will be
permanently archived upon acceptance of the manuscript. The supporting data has been included in the article's supplementary material (Appendix 4).

# Competing interests
We have no competing interests.

# Author's Contributions
CTC, WKC, and JHW conceived and designed the methodology of the
study. CTC, WKC performed the statistical analyses. JMM, REM, and RTK contributed in drafting
the manuscript. All authors provided editorial advice, approved the final version of the manuscript,
and agree to be accountable for all aspects of the work.

@@

```{r Figure 3, echo=FALSE, fig.cap="Reproduced phylogenetic tree (Jetz et al. 2012) of our 477 species (see Appendix 4 for a full list of species with their urbanization indices), demonstrating clustering of species' associations with the urbanization index relative to phylogeny. There are two representations of the urbanization index for each species: (1) coloring ranges from a low urbanization index (dark red) to a high urbanization index (dark blue), with (2) magnitude of the urbanization index (barplot). The urbanization index was the median of the distribution for a species' average radiance values, as shown in Fig. 1. For reference, Figure S3 is a phylogenetic tree with the names of each tip displayed, and we also provide an interactive visualization (https://johnwilshire.github.io/bird_urbanness_web_app/)."}
knitr::include_graphics("finalFigs/Figure3.pdf")
```

@@

```{r Figure S2, echo=FALSE, fig.cap="Correlations between continuous traits, number of records, and number of unique localities used in the analysis."}
knitr::include_graphics("finalFigs/FigureS2.pdf")
```

```{r Figure S3, echo=FALSE, fig.cap="Reference phylogenetic tree of the 477 bird species in the analysis, with labeled tips, based on Jetz et al. (2012). We also provide an interactive visualization for better interpreting the phylogenetic tree (https://johnwilshire.github.io/bird_urbanness_web_app/)"}
knitr::include_graphics("finalFigs/FigureS3.pdf")
```

```{r Figure S4, echo=FALSE, fig.cap="The same phylolm model was fitted to 1000 of the family of trees presented by Jetz et al. (2012) to account for one aspect of phylogenetic uncertainty. Density plots show estimated parameters for the most important traits across models."}
knitr::include_graphics("finalFigs/FigureS4.pdf")
```

>>>>

author: |
  | Corey T. Callaghan^[Corresponding author: c.callaghan@unsw.edu.au] $^1$, Richard E. Major $^1$$^,$$^2$, John H. Wilshire $^1$, John M. Martin $^1$$^,$$^3$, Richard T. Kingsford $^1$, William K. Cornwell $^1$$^,$$^4$
  | $^1$Centre for Ecosystem Science, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia; $^2$Australian Museum Research Institute, Sydney, NSW, Australia; $^3$Royal Botanic Gardens and Domain Trust, Sydney, NSW, Australia; $^4$Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia
date: ""
output:
  rmarkdown::pdf_document:
    fig_caption: yes    
    includes:  
      in_header: preamble-latex.tex
fontsize: 12pt
link-citations: no
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/oikos.csl
bibliography: references.bib
---

@@

\newpage
# INTRODUCTION
Species' specialization, based on a suite of biological traits, can dictate population-level responses to changing environments. Generally, species with limited niche breadth experience higher extinction risk than those with wider niche breadth [@boulangeat2012niche]. Further, some organisms have traits that allow for adaptation of their niches with changing environments, leading to potential ecological inheritance [@odling2013niche]. Urbanization drastically changes natural habitats through destruction, fragmentation, and isolation [@marzluff2001restoration; @melles2003urban; @loram2007urban], often providing ecologically novel landscapes for species. Some species successfully adapt, colonize, and thrive in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation], while others, common in surrounding natural habitats, rarely occur in urban environments [@tratalos2007bird; @evans2011makes; @gatesire2014bird]. Urbanization selects for specific sets of species, based on the regional pool [@aronson2016hierarchical], globally threatening biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects], a major concern for conservation [@czech2000economic; @dearborn2010motivations]. It is important to identify which ecological and life history traits most associate with successful or unsuccessful species in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful; @mcdonnell2015adaptation]. Identifying these traits can improve understanding of relationships between biodiversity and urbanization, and guide protection of the species most at-risk of extinction, through habitat protection, management, and restoration.

Many ecological and life history traits are associated with bird species in urban environments, including degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution]; nesting substrate [@kark2007living; @croci2008does; @conole2011functional]; diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes]; range size [@jokimaki2000artificial; @chace2006urban; @croci2008does]; migratory status [@friesen1995effects]; niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes]; and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific] (see Appendix 1). In Jerusalem, urban bird species were best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while, throughout France and Switzerland, urban bird species were positively associated with forested habitats, sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does]. Despite the breadth of research which has focused on ecological and life history traits associated with urbanization in birds, the results are severely mixed and frequently contradictory. For example, residual brain size in birds is sometimes positively associated with urbanization [@maklakov2011brains; @moller2015brain] and sometimes not [@kark2007living; @evans2011makes]. Similarly, contrasting results have been reported for annual fecundity [cf., @croci2008does; @moller2009successful; @evans2011makes] and niche breadth [cf., @kark2007living; @evans2011makes]. Ultimately, a more comprehensive and generalized understanding of ecological and life history traits associated with urban-tolerant birds is necessary.

Owing to the contrasting and mixed results are methodological shortcomings. Most analyses assign species to discrete a priori groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or strictly presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful]. These approaches severely limit the capacity to generalize, as it assumes grouped species respond equally to urbanization [@lepczyk2008human; @evans2011makes]. Terminology also varies among studies [cf., @kark2007living; @croci2008does; @geschkecompact], affecting interpretation [@mcdonnell2015adaptation; @fischer2015categorizing]. There is clearly a need for continuous measures of urbanization [@evans2011makes], where a species is assessed over its entire distribution in response to urban environments. Further, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does] or temporally limited to a single breeding season [@blair1996land] or even two-month survey period [@kark2007living], restricting the generalizability of the results. Broad-scale empirical data allow hypotheses to be tested at large spatiotemporal scales. Citizen science can provide such data [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen], and can be used to describe patterns of abundance, distribution, and functional composition of birds in urban ecosystems [@mccaffrey2005using; @la2014role; @morelli2016evidence; @lepczyk2017global @callaghan2017assessing].

We aimed to determine what suite of life history traits predicted avian tolerance to urbanization at a continental scale. Our study represented the largest sample size and largest spatial scale ever investigated, with regards to ecological and life history traits and urban-tolerance in birds. We predicted that different traits would positively and negatively associate with urbanization. For generality, our analyses had to (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3) include a broad-range of traits, and (4) employ a continuous measure of urbanization. We integrated a series of large datasets to establish a novel continuous urbanization index to test associations with ecological and life history traits influencing species' responses to urbanization. We first investigated whether our urbanization index was statistically independent in relation to phylogeny (i.e., phylogenetic signal - [@revell2008phylogenetic; @evans2011makes]), given our comprehensive sample size. We then developed models, including and excluding phylogenetic effects, to test for associations between our urbanization index and a suite of continuous and categorical predictor variables representing species' traits. Importantly, given the mixed results of previous research assessing this question and thus competing a priori hypotheses, we found it important to include all predictor variables in the models.

# METHODS
## Bird observation data
We collated bird observations throughout mainland Australia from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists. Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird], often comparable to trained observers [@callaghan2015efficacy]. We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018. This temporal scale corresponded to the richest period of eBird data, minimizing undue leverage of mismatch between changes in response variables over which bird observations were collected.

We filtered bird observations [@callaghan2017assessing; @la2014role], only including observations on complete checklists and which were either travelling, random, stationary, area, or followed BirdLife Australia protocols [@barrett2003new] (see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types). We then filtered the checklists to those which travelled < 5 km or covered < 500 Ha, to ensure the highest quality data were used. Any checklists shared among multiple observers were subsampled, including only one checklist to avoid duplication. We were left with 637,482 eligible checklists throughout Australia. Seabirds (e.g., gannets, petrels, shearwaters) were omitted from analyses as we did not expect any correlation with seabirds using urban areas. Further, only species with > 100 observations were analyzed. In total 580 species, most (~80%) of Australia's bird species, were candidates for analyses. We used differing taxonomic authorities (see Appendix 2).

## Urbanization index
A number of measures of urbanization have previously been used [@sanderson2002human; @hahs2006selecting], but we focused on a continuous measure that is globally applicable (i.e., applicable and easily downloaded for anywhere in the world). As such, we used VIIRS night-time lights [@elvidge2017viirs] as a continuous measure for urban habitat, measuring the electric lighting associated with human settlement (Fig. S1). This has filtered out background noise (e.g., fires, degraded data, and other light source contamination) to ensure validity of the data and its association with human settlement [@elvidge2017viirs]. There is a strong positive relationship between VIIRS night-time lights and human population density (Fig. S2), another commonly used proxy for level of urbanization. For our checklists and their associated species, we used the spatial coordinates and Google Earth Engine [@gorelick2017google], to assign the average radiance value, calculated from the VIIRS layer. For each checklist, the average radiance was calculated within a 5 km buffer, to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols. Any negative radiance values were forced to 0.00001 [*sensu* @ou2015evaluation]. Every bird species subsequently had a distribution of average radiance values (Fig. 1), with the median of each species' distribution becoming the response variable. There was no definitive range for this index, dependent on a species' distribution of observations.

@@

## Statistical analyses
Before modelling of predictor variables (Table 1), we investigated multi-collinearity among continuous variables (Fig. S3), using the corrplot package [@corrplot2017]. We used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation among each other; all traits < 2 suggested negligible correlation (Supplementary Table 1). We used two independent linear modelling approaches: one without and one with phylogenetic effects. Phylogenetic effects are defined as the tendency of closely related organisms to have similar life history components due to their shared evolutionary history [@derrickson1988taxon; @mckitrick1993phylogenetic]. Both approaches were used because of shortcomings [@cadotte2017phylogenies] and importance [@losos2008phylogenetic; @graham2018phylogenetic] of incorporating evolutionary distinctness in ecological research.

We performed linear regression models with the response variable being the species-specific urbanization index and the predictor variables the 22 ecological and life history traits. This was done without and with controlling for phylogenetic effects (see below). All linear regression models were weighted [@solon2015weighting], using a source of error associated with our urbanization indices. This error was the proportion of the number of total observations for a species, divided by the number of its unique localities (Fig. S3 shows the correlation between the two). This weighted the model-fitting to those species with more observations in more unique locations, likely to be the most reliable urbanization indices. This approach also addressed biases in sampling of eBird data, where some birds have a high number of observations from a small number of locations such as known roosting sites [e.g., @callaghan2018avian]. All models and analyses were conducted in R [@r2017coreteam], relying on the tidyverse workflow [@wickham2017tidyverse]. Statistical significance was concluded as $\alpha$ < 0.05. Variability of means is reported as standard deviations.

@@

## Phylogenetic analyses
### Phylogenetic signal
There was a distinct relationship between the phylogenetic signal and our urbanization index (Fig. 3; Fig. S4; Table 2), significantly different from random for 4 out of 5 indices (Cmean, I, K.star, and Lambda). For example, the urbanization index was disproportionately important with respect to the phylogenetic tree (Fig. 3), evidenced by the highest urbanization indices clustered throughout the waterbird clade. Further, within a specific clade, there were some species with substantially higher urbanization indices (e.g., *Manorina melanocephala*, Noisy Miner, cf. honeyeaters). We generally found weaker signals for behavioral, compared with morphological traits. For behavioural traits, the estimates of phylogenetic signal were weaker than Brownian motion, but usually significant (Table 2). Additionally, there was a strong phylogenetic signal for body size and brain residual size (Table 2).

@@

# DISCUSSION
The most urban-tolerant species in our analyses, defined based on their urbanization indices, had large niche breadths, reflected in generalism. Conversely, more specialized species were the least urban-tolerant: specialized feeding strategies (i.e., granivorous and insectivorous diets) seldom occurred in urban environments. These results were robust across separate non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig. S5), indicating that urbanization is generally excluding specialist species [@sorace2009distribution; @concepcion2015impacts]. Further, the non-random distribution of the urbanization index across the phylogeny (Fig. 3 & Table 2) is consistent with reduced phylogenetic diversity in urban areas [@la2018phylogenetic]. Particular phylogenetic effects [@derrickson1988taxon; @mckitrick1993phylogenetic] probably allow specific subsets of species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].

@@

In conclusion, our novel continuous urbanization index showed relationships between ecological and life history traits with avian tolerance to urban environments. Our results are also generalizable and comprehensive, given the largest sample size and spatial scale to be used to address this question. Further, the methodology used is globally applicable, dependent only on observational data and trait data, combined with remotely-sensed products. Improving understanding of the temporal nature of these relationships will be important, but we highlight the importance of considering these questions using a continuous measure of urbanization [@evans2011makes] as well as using a phylogenetically informed approach. It is the generalist species [@shochat2006patterns], with high clutch sizes and large residual brain sizes, which are best prepared to forage and survive in the ecologically novel aspects of urban environments.

# Data archiving
This work relied heavily on open-source, fully accessible datasets.
The necessary portions of those datasets, along with code to reproduce the analyses are available at
a github repository, and will be
permanently archived upon acceptance of the manuscript. The supporting data has been included in the article's supplementary material (Appendix 4).

@@

```{r Figure 3, echo=FALSE, fig.cap="Reproduced phylogenetic tree (Jetz et al. 2012) of our 477 species (see Appendix 4 for a full list of species with their urbanization indices), demonstrating clustering of species' associations with the urbanization index relative to phylogeny. There are two representations of the urbanization index for each species: (1) coloring ranges from a low urbanization index (dark red) to a high urbanization index (dark blue), with (2) magnitude of the urbanization index (barplot). The urbanization index was the median of the distribution for a species' average radiance values, as shown in Fig. 1. For reference, Figure S3 is a phylogenetic tree with the names of each tip displayed."}
knitr::include_graphics("finalFigs/Figure3.pdf")
```

@@

```{r Figure S2, echo=FALSE, fig.cap="The relationship between VIIRS night-time lights and human population density, fitted with a linear model (blue line)."}
knitr::include_graphics("finalFigs/FigureS2.pdf")
```

```{r Figure S3, echo=FALSE, fig.cap="Correlations between continuous traits, number of records, and number of unique localities used in the analysis."}
knitr::include_graphics("finalFigs/FigureS3.pdf")
```

```{r Figure S4, echo=FALSE, fig.cap="Reference phylogenetic tree of the 477 bird species in the analysis, with labeled tips, based on Jetz et al. (2012). We also provide an interactive visualization for better interpreting the phylogenetic tree (https://johnwilshire.github.io/bird_urbanness_web_app/)"}
knitr::include_graphics("finalFigs/FigureS4.pdf")
```

```{r Figure S5, echo=FALSE, fig.cap="The same phylolm model was fitted to 1000 of the family of trees presented by Jetz et al. (2012) to account for one aspect of phylogenetic uncertainty. Density plots show estimated parameters for the most important traits across models."}
knitr::include_graphics("finalFigs/FigureS5.pdf")
```
